---
title: "Covid91 vaccine study Final2024"
author: "Daniel Purri!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the Covid vaccine effetive in males?

$H_0$: The new preventative vaccine (Covid921) does not exhibit any effectivness on males.  
$H_A$: The new preventative vaccine (Covid921) does exhibit effectivness on male.  

Cat~Cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```
This graph shows that the vaccine is effective in preventing covid19 in males because the amount of males who received the vaccine and contracted covid19 was lower than the men who had received the placebo.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

The table shows 26.20% of men received the vaccine and had covid19 while 73.80% of men received the placebo got covid 19.

##Inferential Results  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

So, the p-value from the test is 0.00000000000000022, which is less than 0.05, we have to reject the null hypothesis. The odds ratio from the fisher's exact test was 2.815 and this has a 2.8 chance of being effective in men



## Females


##Methods  
$H_0$: The preventative vaccine does not exhibit its effective in females        
$H_A$: The preventative vaccine exhibits its effective in females.  
Cat~Cat. 

```{r}
Females = FinalData[FinalData$sex=='F',]
head(Females)
```

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

The bar graph shows that the vaccine is more effective in women that got covid19 compared to women who got the placebo

##Numerical Results  
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The numerical results shows that something close to 35.07% of women that got the vaccine had covid19, but the women who got covid19 and also got the placebo which was 64.93%. Shows that the vaccine is good

##Inferential Results    
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the test is 0.00000001914 adn is less than 0.05 so, we reject the null hypothesis. The odds ratio from the fisher's exact test is 1.856 which means that there is approximately a 1.86 chance that the vaccine will be effective in women


## LGBTQ

##Methods  
$H_0$ The preventative vaccine does not exhibit its effective in people who are LGBTQ.        
$H_A$ The preventative vaccine exhibits its effective in people who are LGBTQ.    
Cat~Cat

```{r}
gay = FinalData[FinalData$LGBTQ=='gay',]
head(gay)
```

##Graphical Results  

```{r}
barchartGC(~infected + treatment,data=gay)
barchartGC(~infected + treatment,data=gay, type="percent")
```

The bar graph shows that the vaccine is good in preventing covid19 in people who are LGBTQ 

##Numerical Results  
```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The table shows that 20% of LGBTQ people that got the vaccine also got covid19. Compared to the LGBTQ people who received the placebo and contracted vaccine which is 80%. This means that the vaccine is good

##Infrential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the test is 0.000002266 which is less than 0.05 so, we reject the null hypothesis. The odds ratio from the fisher's exact test is 4.27 which means that the vaccine has a 4.27 chance of being good.

## Druggies

##Methods  
$H_0$ The preventative vaccine does not exhibit its effective in people who are drug users.    
$H_A$ The preventative vaccine exhibits its effective in people who are drug users.  
Cat~Cat

```{r}
DrugUser = FinalData[FinalData$DrugUser=='yes',]
```

##Graphical Results    
```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```

The bar graphs shows that vaccine is more likely to be less effective in people who are drug users

##Numerical Results     
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The table shows that 91.67% of drug users that got covid19 also got the vaccine. So this shows that the vaccine is not relly good and effective with drugs

##Inferential Results      
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the chi-squared test is 0.0000000001414 which is less than 0.05 so, we reject the null hypothesis. Pretty much vaccine has not much effectiveness.

# Overall Results and Conclusions

##Males     

Based on the data we got, it looks like the vaccine actually works for guys. The table shows that only 26.2% of men who got the vaccine ended up getting covid, while 73.8% of the men who got the placebo caught it. That’s a big difference. The p-value was like 0.00000000000000022, which is way smaller than 0.05, so yeah—we reject the null. That means there’s solid evidence the vaccine does something. The odds came out to around 2.8 times more effective, so we can say the vaccine is pretty reliable for preventing covid in men.  


##Females   

From the data we got, we can say the vaccine works for women too. The table shows that 35.07% of the women who got covid had the vaccine, and 64.93% were the ones who got the placebo. That’s a big enough gap. The p-value was 0.00000001914, which is definitely less than 0.05, so we reject the null again. That means there’s strong evidence the vaccine helps. The odds came out to like 1.86, so yeah, we can say the vaccine seems to help prevent covid in females too.     


##LGBTQ    

Looking at the data we got, it shows the vaccine works pretty well for LGBTQ folks too. The table shows that only 20% of LGBTQ people who got covid had the vaccine, while 80% of the ones who got the placebo ended up with it. So yeah, that’s a big difference and shows the vaccine is doing its job. The p-value was 0.000002266, which is way under 0.05, so we reject the null again. That means there’s strong evidence here. The odds ratio from Fisher’s exact test came out to 4.27, which basically means the vaccine has over a 4 times better chance at helping prevent covid in LGBTQ people. So yeah, it’s pretty safe to say the vaccine works for this group too.

##Drug Users  

From the data we got, it looks like the vaccine doesn’t really work well for people who use drugs. The table shows that 91.67% of drug users who got covid had the vaccine, and only 8.33% got the placebo, which kinda flips everything. The p-value was 0.0000000001414, so yeah, we reject the null again, but this time it’s showing something different. The odds ratio from Fisher’s test was 0.087, which means the vaccine is actually way less likely to help drug users. So based on this, we can say the vaccine probably won’t be that effective at preventing covid in this group.