---
title: "Covid91 vaccine study Final2024"
author: "Daniel Purri!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the Covid vaccine effetive in males?

$H_0$: The new preventative vaccine (Covid921) does not exhibit any effectivness on males.  
$H_A$: The new preventative vaccine (Covid921) does exhibit effectivness on male.  

Cat~Cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```
This graph shows that the vaccine is effective in preventing covid19 in males because the amount of males who received the vaccine and contracted covid19 was lower than the men who had received the placebo.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

The table shows 26.20% of men received the vaccine and had covid19 while 73.80% of men received the placebo got covid 19.

##Inferential Results  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

So, the p-value from the test is 0.00000000000000022, which is less than 0.05, we have to reject the null hypothesis. The odds ratio from the fisher's exact test was 2.815 and this has a 2.8 chance of being effective in men



## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions