---
title: "Covid91 vaccine study Final2024"
author: "Daniel Purri!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

Is the Covid vaccine effetive in males?

$H_0$: The new preventative vaccine (Covid921) does not exhibit any effectivness on males.  
$H_A$: The new preventative vaccine (Covid921) does exhibit effectivness on male.  

Cat~Cat

```{r}
Males = FinalData[FinalData$sex=='M',]
head(Males)
```

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Males)
barchartGC(~infected + treatment,data=Males, type="percent")
```
This graph shows that the vaccine is effective in preventing covid19 in males because the amount of males who received the vaccine and contracted covid19 was lower than the men who had received the placebo.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```

The table shows 26.20% of men received the vaccine and had covid19 while 73.80% of men received the placebo got covid 19.

##Inferential Results  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

So, the p-value from the test is 0.00000000000000022, which is less than 0.05, we have to reject the null hypothesis. The odds ratio from the fisher's exact test was 2.815 and this has a 2.8 chance of being effective in men



## Females


##Methods  
$H_0$: The preventative vaccine does not exhibit its effective in females        
$H_A$: The preventative vaccine exhibits its effective in females.  
Cat~Cat. 

```{r}
Females = FinalData[FinalData$sex=='F',]
head(Females)
```

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")
```

The bar graph shows that the vaccine is more effective in women that got covid19 compared to women who got the placebo

##Numerical Results  
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

The numerical results shows that something close to 35.07% of women that got the vaccine had covid19, but the women who got covid19 and also got the placebo which was 64.93%. Shows that the vaccine is good

##Inferential Results    
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the test is 0.00000001914 adn is less than 0.05 so, we reject the null hypothesis. The odds ratio from the fisher's exact test is 1.856 which means that there is approximately a 1.86 chance that the vaccine will be effective in women


## LGBTQ

##Methods  
$H_0$ The preventative vaccine does not exhibit its effective in people who are LGBTQ.        
$H_A$ The preventative vaccine exhibits its effective in people who are LGBTQ.    
Cat~Cat

```{r}
gay = FinalData[FinalData$LGBTQ=='gay',]
head(gay)
```

##Graphical Results  

```{r}
barchartGC(~infected + treatment,data=gay)
barchartGC(~infected + treatment,data=gay, type="percent")
```

The bar graph shows that the vaccine is good in preventing covid19 in people who are LGBTQ 

##Numerical Results  
```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The table shows that 20% of LGBTQ people that got the vaccine also got covid19. Compared to the LGBTQ people who received the placebo and contracted vaccine which is 80%. This means that the vaccine is good

##Infrential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the test is 0.000002266 which is less than 0.05 so, we reject the null hypothesis. The odds ratio from the fisher's exact test is 4.27 which means that the vaccine has a 4.27 chance of being good.

## Druggies

##Methods  
$H_0$ The preventative vaccine does not exhibit its effective in people who are drug users.    
$H_A$ The preventative vaccine exhibits its effective in people who are drug users.  
Cat~Cat

```{r}
DrugUser = FinalData[FinalData$DrugUser=='yes',]
```

##Graphical Results    
```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")
```

The bar graphs shows that vaccine is more likely to be less effective in people who are drug users

##Numerical Results     
```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The table shows that 91.67% of drug users that got covid19 also got the vaccine. So this shows that the vaccine is not relly good and effective with drugs

##Inferential Results      
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value from the chi-squared test is 0.0000000001414 which is less than 0.05 so, we reject the null hypothesis. Pretty much vaccine has not much effectiveness.

# Overall Results and Conclusions